Universal Anonymous Signatures

### 2022-04-02 prism-backend/research/gsac/introduction.tex
Merged GSAC and UAS into uas.tex paper. Defining concrete instantiation of GSAC.

In this work, we combine two closely related schemes -- group signatures (GS)
and anonymous credentials (AC) -- which, for some reason, have followed a
separate path up to now, beyond individual constructions from both fields
drawing inspiration (and techniques) from one another. This combination is in
some sense a natural evolution, which we do believe may have relevant impact in
practical settings. On the one hand, anonymous credentials enable a richer
functionality in regards to what one can prove about his own identity, in a
privacy-preserving way. Academic proof of this is the fact that AC systems have
naturally evolved to enabling arbitrary claims that only reveal the minimally
possible 1-bit of information \needcite. Also, AC systems have been proposed to
go even beyond what one can prove about his identity, covering also orthogonal
functionality such as delegation \needcite, or aggregation of multiple
credentials into one \needcite. Even outside academia, prominent
technical workgroups \needcite, startups \needcite, and big techs \needcite are
promoting and adopting solutions which are directly inspired by AC systems. On
the other hand, GS 
schemes have inspection functionality (e.g., opening and linking of signatures)
at their core, which is reflected formally through non-frameability properties.
This has originated a large body of research focusing on the construction and
formalisation of variants of these notions \needcite. Consequently, group
signatures tend to be proposed for scenarios that require some sort of
de-anonymization \needcite. Indeed, this is not possible in ACs, which at most
support revocation of issued credentials, which essentially can compare to
linking, but not to opening-related notions.

Our goal is to get the best of both worlds: leverage the richer functionality
of ACs in what respects to proving, in a privacy-preserving manner, useful
claims on one's identity; and, at the same time support, through inspection
capabilities, use cases that may require some sort of de-anonymiztion. We
could target the ideal case of allowing to create proofs of arbitrary predicates
on one's identity (as some AC systems do), as well as provide flexible or several
notions of non-frameability. However, this being -- as far as we know -- the
first work that explores this intersection, we opt for keeping things simple
for the sake of facilitating technical precission and readability, and fostering
discussion. Therefore, we focus on the case of selective attribute disclosure
for ACs, and conventional opening for GSs, building a GSAC scheme that supports
both.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "sok-privsig"
%%% End:

>>>>

\commentwho{Jesus}{The following will most probably change completely if we
  are to submit this for publication (it least, it probably needs some deep
  reorganisation). However, this should capture tha main ideas.}

In this work, we focus on the generalization of privacy-preserving cryptographic
systems for authentication and identification. Two major research lines within
cryptography that work in this space are group signatures (GS) and anonymous
credentials (AC). In fact, despite having followed separate paths, the overall
approaches to build them are quite similar in a general sense. Consequently,
both have fed from each other.
%
In short, GSs allow a group of users, accepted into the group by an issuer, to
create signatures on behalf of the group, without revealing their identity. Only
an opener/inspector can (sometimes partially) deanonymize these signatures. In
ACs, users obtain credentials, typically containing multiple attributes, from
trusted issuers. Subsequently, users can prove ownership of these credentials,
usually revealing only subsets (or arbitrary claims) of these attritubes, while
keeping the rest hidden.

On the one hand, the security models of GS and AC hold important similarities.
Security in group signatures is typically captured with game-based definitions,
where the main properties are anonymity, traceability and non-frameability.
Anonymity ensures privacy, essentially requiring that group signatures do not
leak information about their signer. Traceability and non-frameability are
unforgeability-related properties. Traceability requires that any valid group
signature must have been created by a user who joined the group;
non-frameability requires that no adversary can create a signature that somehow
``blames'' an honest user. Of course, variations in the functionality of group
signatures frequently require specialized properties, but these three are always
the main ones. In anonymous credentials, security is also usually captured with
game-based definitions (although UC models also exist). But, for anonymous
credentials, the main properties (again, up to variations) are simply anonymity
and unforgeability. Anonymity captures a similar privacy notion than group
signatures. Unforgeability of anonymous credentials also maps more clearly to
traceability in group signatures (non-frameability assumes a corrupt issuer,
which must be honest in traceability and unforgeability), and requires that
no adversary can succeed in proving ownership of a credential it does not own.
In both cases -- anonymity and unforgeability in anonymous credentials -- the
focus is on preventing leakage (anonymity) or forging (unforgeability) of the
attributes in a credential. Since group signatures do not (usually) have
attributes, this is a core difference in the modelling.

The first thing that raises attention here is the fact that, despite following
very similar approaches to build them, there is no equivalent property for
non-frameability in anonymous credentials. The most probable reason is that
anonymous credentials do not include the concept of ``opening'' which, in
group signatures, allows to extract the identity of the signer and is essential
to the non-frameability definition\footnote{Exceptions exist \cite{dl21}, but
  the definition is much easier with opening.}. Something similar (more limited
than opening, though) is sometimes considered by means of revocation of
anonymous credentials -- revoking credentials is essentially equivalent to
being able to link signatures by the same signer, which is considered in some
variants of group signatures \needcite. However, while revocation may be enough
if we aim to use credentials ``just'' for authentication, if we target more
advanced use cases, we may need to learn something else from the credential (or
its owner), beyond whether or not it has been revoked. Then, without an
equivalent to non-frameability, we would not be able to make any security claim
in the presence of dishonest issuers. On the other hand, it is
also noteworthy that the literature of group signatures has barely explored
the inclusion of attributes. With the exception of a few works on Attribute-Based
Group Signatures (ABGS) \needcite, members of a group in group signatures just
get a credential attesting for their membership to the group. Moreover, the
existence of attributes in anonymous credentials would seem to have been
instrumental in the surge of real-world use cases in the past few years. Indeed,
initiatives such as W3C's Verifiable Credentials (VCs) \needcite, or large
projects such as Idemix \needcite, which has in turn influenced VCs, are heavily
based on concepts from the anonymous credential literature, even to the point
that Idemix and its ``blockchain-based successor'', Indy \needcite, are now a
type of verifiable credentials \needcite.

Following the previous reasoning, we identify two aspects in which we can
improve group signatures and anonymous credentials. The first is by coming up
with some sort of primitive that combines both lines. That is, a scheme which
supports both (a variant of) opening and non-frameability as well as
attribute-based credentials. This seemingly natural evolution would thus
enable rich use cases in which accountable (by means of opening) and expressive
(by means of the attributes) authentication is necessary. While this is in
line with the previously mentioned ABGS works, we argue that our approach
is more general and flexible. Still, we anticipate that this may not be enough
to reach a scheme that can adapt to the majority of situations that may arise in
the real world. The reason is that the issuance and opening capabilities in the
previous schemes are generally either too rigid or have lacked a formal analysis
so far. For instance, in group signatures, the issuance protocol has barely been
the focus of study. Essentially, this interactive protocol requires the
prospective member to prove, in zero-knowledge, knowledge of a secret key
that will be attested in the membership credential. On the other hand, multiple
variants of the issuance protocol exist in anonymous credentials: protocols that
assume that the issuer knows all the attributes to be included in the credential,
protocols that allow blind issuance of attributes, protocols that let the user
leverage attributes in previous credentials, protocols that include ad hoc
rules at issuance time to enable extra functionality, such as delegation
\todo{cross-check that all this, especially the latter, is correct; and add
  cites}, etc. The problem is that a single change in this aspect, requires to
completely rethink the model. Something similar happens with the inspection
capabilities in group signatures. This frequently takes the shape of opening,
but many variants also exist, message-depdendent opening, linking, or variants
of linking, such as restricted linking only by other members of the group,
user controlled linking, or sequential linking \needcite. Again, any small
change in the inspection capabilities requires a full redesign of the model.
This rigidness in inspection is specially crucial, as inspection is the key
to the very nature of the privacy notion in group signatures. Taking this into
account, it is understandable that many have tried to find some sort of
equilibrium between full anonymity (no inspection at all), and full
accountability (by means of conventional open). Thus, we aim at finding a
generic way to support multiple variants of inspection, from the common model.
Finally, actual evaluation of group signatures is the last component that is
too rigid. Namely, in group signatures, when verifying a signature, either it
has been rightly produced by a group member using his secret key and membership
credential, or not. Luckily for us, anonymous credentials have done a great job
in supporting high expressiveness into what a presentation can tell beyond just
proving knowledge of a (group) key. Thus, by integreating anonymous credentials
and group signatures in a sort of combined primitive, we get \emph{for free}
the benefits of this high expressiveness that previous work on anonymous
credentials have reached in regards to what one can prove given a credential.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "uas"
%%% End:

### 2022-04-02 prism-backend/research/gsac/model-uas.tex
Merged GSAC and UAS into uas.tex paper. Defining concrete instantiation of GSAC.

\section{Model for \UAS Schemes}
\label{sec:model-uas}

@@

\subsection{Syntax}
\label{ssec:syntax}

@@

\begin{description}
\item[$\parm \gets \Setup(\secpar)$.] Given a security parameter \secpar,
  returns a global system parameter variable \parm.
\item[$(\ipk,\isk) \gets \IKeyGen(\parm,\fissue)$.] Given global system
  parameters \parm, and the function \fissue to be used to check that credential
  requestors meet the conditions to be issued a credential, an issuer runs
  \IKeyGen to generate its issuing key pair. Hereafter, we assume that the
  public part \ipk is added to the group public key \gpk, as well as \fissue.
\item[$(\opk,\osk) \gets \OKeyGen(\parm,\finsp)$.] Given global system
  parameters \parm, and function \finsp, an inspector runs \OKeyGen to generate
  its inspecting key pair. The function \finsp defines the type of utility that
  will be extractable from those pairs that do not meet conditions defined
  at signing time (typically, by verifiers). Hereafter, we assume that the
  public part \opk is added to the group public key \gpk, along with \finsp.
\item[$(\upk,\usk) \gets \UKeyGen(\parm)$.] Given global system parameters
  \parm, returns a user's key pair.
\item[$\langle \cred/\bot,\utrans/\bot \rangle \gets
  \langle
  \Obtain(\usk,\attrs,\ldblbrace (gpk_i,\cred_i)\rdblbrace_{i \in \Issuers}),
  \Issue(\isk,\upk,\attrs,\ldblbrace \gpk_i) \rdblbrace_{i \in \Issuers})
  \rangle$.] %
  This interactive protocol lets a user with key pair (\upk,\usk) running the
  \Obtain process, obtain a credential \cred from an issuer in the system, on
  attribute set $\attrs$ plus, possibly, blinded attributes attested by
  previously obtained credentials $\cred_i$, from a multiset of other issuers
  (which may include the one to whom a new credential is being requested) with
  public keys $\gpk_i$. The user outputs the produced credential \cred, while
  the issuer outputs the protocol transcript \utrans for the produced
  credential. To ease notation, we assume that the group to which each
  credential belongs is available from the credential itself, and therefore
  omit the $\gpk_i$ values from the syntax, unless necessary to avoid ambiguity.
\item[$\sig \gets \Sign(\gpk,\usk,\cred,\msg,\feval)$.]
  The user with with secret key \usk and credential \cred, produces a signature
  \sig over message \msg, meeting the conditions defined by \feval.
\item[$1/0 \gets \Verify(\gpk,\sig,\msg,\feval)$.] Checks whether \sig is a
  valid signature, over message \msg, from a user in the group with public key
  \gpk, for signing function \feval.
\item[$(\y,\iproof)/\bot \gets \Inspect(\gpk,\osk,\trans,\sig,\msg,\feval)$.] %
  Executed by the inspector with private key \osk. Receives a signature \sig
  over message \msg and evaluation function \feval. If \trans contains a set of
  valid transcripts corresponding
  to the $\langle\Obtain,\Issue\rangle$ interactive protocol execution that
  issued the credential used to produce \sig, the function outputs a value $\y$
  derived from the signer's public key, credential, message, and signature, as
  well as a proof of correct inspection. We sometimes abuse notation, and write
  $\trans[\uid]$ to mean that the \Inspect function is operating only for the
  obtain transcripts related to user \uid.
\item[$1/0 \gets \Judge(\gpk,\y,\iproof,\sig,\msg,\feval)$.] %
  Checks if \iproof is a valid inspection correctness proof for the value \y,
  obtained by applying \Inspect to the the signature \sig over message \msg. 
\end{description}

@@

\begin{description}
\item[$\fissue: (\upk, \attrs,\scred) \rightarrow 0/1$.] Defined by issuers,
  governs what customized conditions an issuer requires in order to issue
  credentials, when receiving a request from user with public key \upk, for
  attributes \attrs. \fissue may run checks on a (possibly empty) set of
  additional credentials \scred, which may further be used for blind issuance
  \commentwho{Jesus}{can it?}. \fissue returns $1$ to accept a request, $0$ to
  reject it.  
\item[$\feval: (\upk,\cred,\msg) \rightarrow \varrngfeval$.] Can be defined by any
  party, although we anticipate that this will typically be done by either
  signers, verifiers, or some governance organization. It defines customized
  conditions to be met by the owner of a user key \upk and credential \cred,
  in order to sign a concrete message \msg. Its outputs  $\varrngfeval$ must
  belong in a well defined set \rngfeval, with the restriction that it must
  return $0$ when the conditions for signing are not met.
\item[$\finsp: (\varrngfeval,\upk,\cred,\msg) \rightarrow \varrngfinsp$.]
  Defines what utility value, derived from the user's public key, credential,
  and message should be extractable by the group inspector. Note that \finsp
  also receives as input a value in the range of the \feval function, \rngfeval.
  This will allow inspection logic to depend on the value produced by the
  evaluation function. The output of \finsp is a value \varrngfinsp, which must
  belong in a well defined set \rngfinsp.
\end{description}

@@

\subsection{Security Properties}
\label{ssec:security}

@@

\subsection{Functionality ``Transformations'' and Special Cases}
\label{ssec:transformations}

\commentwho{Jesus}{This section is aimed at showing the ``universality'' of
  our model. If I'm not mistaken, ideally, it should support the following
  variations (except the transformation to interactive presentations,
  probably).}

\subsubsection{Transformation to Interactive Presentations}
\label{sssec:interactivetransform}

\todo{Is it possible to give it in some generic sense?}

\subsubsection{Combining Multiple Credentials in One Presentation}

\todo{I think this might be possible if we can use \fissue so that any user can
  create arbitrary groups, where he is the issuer, an can thus run
  \Obtain,\Issue protocols to combine credentials from other groups. This is
  probably inefficient (or, at least, more than doing so wihtout having to
  create ad hoc groups), but at least allows the enhanced functionality...}

\subsubsection{Restricting to Conventional GS and AC Schemes}

\todo{It would be nice to prove that a \GSAC scheme to which we remove the
  \Open/\Judge functions becomes an AC scheme. And conversely, a \GSAC scheme
  where all credentials have no attributes, and where we restrict to only
  one credential per user, becomes a conventional GS scheme.}

\subsubsection{Group Signatures with Message Dependent Opening}

\todo{Describe what type of \finsp would be needed to achieve a functionality
  similar to that of GS-MDO. Does this fit our model straight away? Maybe also
  show how other types of GS could be ``mimicked'' (e.g., GS only with linking).
  This would also be good to test if the model is as generic as I think it is}

\subsubsection{Ring Signatures}

\todo{Can we come up with a (\fissue,\feval,\finsp) tuple that allows us to
  somehow mimic ring signatures (i.e., ``a group signature without issuer nor
  opening''? Does the model support this?}

>>>>

\subsection{Model for \UAS Schemes}
\label{ssec:model-uas}

@@

\subsubsection{Syntax}
\label{sssec:syntax}

@@

\begin{description}
\item[$\parm \gets \Setup(\secpar)$.] Given a security parameter \secpar,
  returns a global system parameter variable \parm.
\item[$(\ipk,\isk) \gets \IKeyGen(\parm,\fissue)$.] Given global system
  parameters \parm, and the function \fissue to be used to check that credential
  requestors meet the conditions to be issued a credential, an issuer runs
  \IKeyGen to generate its issuing key pair. Hereafter, we assume that the
  public part \ipk is added to the group public key \gpk, as well as \fissue.
\item[$(\opk,\osk) \gets \OKeyGen(\parm,\finsp)$.] Given global system
  parameters \parm, and function \finsp, an inspector runs \OKeyGen to generate
  its inspecting key pair. The function \finsp defines the type of utility that
  will be extractable from those pairs that do not meet conditions defined
  at signing time (typically, by verifiers). Hereafter, we assume that the
  public part \opk is added to the group public key \gpk, along with \finsp.
\item[$(\upk,\usk) \gets \UKeyGen(\parm)$.] Given global system parameters
  \parm, returns a user's key pair.
\item[$\langle \cred/\bot,\utrans/\bot \rangle \gets
  \langle
  \Obtain(\usk,\attrs,\ldblbrace (gpk_i,\cred_i)\rdblbrace_{i \in \Issuers}),
  \Issue(\isk,\upk,\attrs,\ldblbrace \gpk_i) \rdblbrace_{i \in \Issuers})
  \rangle$.] %
  This interactive protocol lets a user with key pair (\upk,\usk) running the
  \Obtain process, obtain a credential \cred from an issuer in the system, on
  attribute set $\attrs$ plus, possibly, blinded attributes attested by
  previously obtained credentials $\cred_i$, from a multiset of other issuers
  (which may include the one to whom a new credential is being requested) with
  public keys $\gpk_i$. The user outputs the produced credential \cred, while
  the issuer outputs the protocol transcript \utrans for the produced
  credential. To ease notation, we assume that the group to which each
  credential belongs is available from the credential itself, and therefore
  omit the $\gpk_i$ values from the syntax, unless necessary to avoid ambiguity.
\item[$\sig \gets \Sign(\gpk,\usk,\cred,\msg,\feval)$.]
  The user with with secret key \usk and credential \cred, produces a signature
  \sig over message \msg, meeting the conditions defined by \feval.
\item[$1/0 \gets \Verify(\gpk,\sig,\msg,\feval)$.] Checks whether \sig is a
  valid signature, over message \msg, from a user in the group with public key
  \gpk, for signing function \feval.
\item[$(\y,\iproof)/\bot \gets \Inspect(\gpk,\osk,\trans,\sig,\msg,\feval)$.] %
  \todo{I think we do not need \trans here, see the generic construction. (This
    was somehow inherited from \cite{bsz05}.) However, if we do remove it, we'd
    need to redefine the traceability game -- it'd need to depend on \Identify,
    which kind of makes sense too.}
  Executed by the inspector with private key \osk. Receives a signature \sig
  over message \msg and evaluation function \feval. If \trans contains a set of
  valid transcripts corresponding 
  to the $\langle\Obtain,\Issue\rangle$ interactive protocol execution that
  issued the credential used to produce \sig, the function outputs a value $\y$
  derived from the signer's public key, credential, message, and signature, as
  well as a proof of correct inspection. We sometimes abuse notation, and write
  $\trans[\uid]$ to mean that the \Inspect function is operating only for the
  obtain transcripts related to user \uid.
\item[$1/0 \gets \Judge(\gpk,\y,\iproof,\sig,\msg,\feval)$.] %
  Checks if \iproof is a valid inspection correctness proof for the value \y,
  obtained by applying \Inspect to the the signature \sig over message \msg. 
\end{description}

@@

\begin{description}
\item[$\fissue: (\upk, \attrs,\scred) \rightarrow 0/1$.] Defined by issuers,
  governs what customized conditions an issuer requires in order to issue
  credentials, when receiving a request from user with public key \upk, for
  attributes \attrs. \fissue may run checks on a (possibly empty) set of
  additional credentials \scred, all bound to \upk, which may further be used
  for blind issuance \commentwho{Jesus}{can it?}. \fissue returns $1$ to accept
  a request, $0$ to reject it.  
\item[$\feval: (\upk,\cred,\msg) \rightarrow \varrngfeval$.] Can be defined by
  any party, although we anticipate that this will typically be done by either
  signers, verifiers, or some governance organization. It defines customized
  conditions to be met by the owner of a user key \upk and credential \cred,
  bound to \upk, in order to sign a concrete message \msg. Its outputs
  $\varrngfeval$ must belong in a well defined set \rngfeval, with the
  restriction that it must return $0$ when the conditions for signing are not
  met.
\item[$\finsp: (\varrngfeval,\upk,\cred,\msg) \rightarrow \varrngfinsp$.]
  Defines what utility value, derived from the user's public key, credential,
  and message should be extractable by the group inspector. Note that \finsp
  also receives as input a value in the range of the \feval function, \rngfeval.
  This will allow inspection logic to depend on the value produced by the
  evaluation function. The output of \finsp is a value \varrngfinsp, which must
  belong in a well defined set \rngfinsp.
\end{description}

@@

\subsubsection{Security Properties}
\label{sssec:security}

@@
